Tracing the cultural evolution of meaning at the semantics-pragmatics interface

### 2016-12-29 paper/new_simulation_setup.tex
new simulation specs update

\begin{table}
  \centering
\begin{center}
  \begin{tabular}{lccccl}
    \toprule
    intuitive name
    & \snone
    & \ssome
    & \sall
    & min.~derivation length
    & formula
    \\ \midrule
    ``all''
    & 0
    & 0
    & 1
    & $3$
    & $A \subseteq B$
    \\
    ``some but not all''
    & 0
    & 1
    & 0
    & $7$
    & $A \cap B \neq \emptyset \wedge A \neq \emptyset$
    \\    
    ``some''
    & 0
    & 1
    & 1
    & $4$
    & $A \cap B \neq \emptyset$
    \\
    ``none''
    & 1
    & 0
    & 0
    & $4$
    & $A \cap B = \emptyset$
    \\
    ``none or all''
    & 1
    & 0
    & 1
    & $8$
    & $\neg(A \cap B \neq \emptyset \wedge A \neq \emptyset)$
    \\
    ``not all''
    & 1
    & 1
    & 0
    & $4$
    & $\neg (A \subseteq B)$
    \\
    \bottomrule
  \end{tabular}
\end{center}
\caption{Available concepts and their minimal derivation length}
\label{tab:concepts}
\end{table}

Identifying three world states requires three messages in $\Messgs$. As before, we call a
lexicon $L$ a mapping $\Messgs \rightarrow C$ from messages to concepts. There are $6^3 = 343$
possible lexica, but many of these assign the same concept to more than one message. For
simplicity, attention could be restricted to only those lexica which avoid expressive
redundancy. Hardcoding such an \emph{mutual exclusivity bias}
\citep[e.g.][]{Clark2009:Lexical-Meaning}, there are 20 non-redundant
lexica.\footnote{Functional pressure towards efficient communication will evidently favor
  non-redundant lexica, so this restriction is fairly innocuous but practical.}

@@

\begin{table}
  \centering
  \begin{align*}
    C & \rightarrow C \wedge C 
    & 
    X & \rightarrow \set{A,B} \\
    C & \rightarrow\neg C 
    & 
    X & \rightarrow X \cap X \\
    C & \rightarrow X \subseteq X
    & 
    X & \rightarrow X \cup X \\
    C & \rightarrow X \neq \emptyset
    & 
    X & \rightarrow X \setminus X \\
    C & \rightarrow X = \emptyset     
    & 
    X & \rightarrow \overline{M}
  \end{align*}
  \caption{Toy grammar in a set-theoretic "language of thought"}
  \label{tab:grammar}
\end{table}

The prior probability of a type is just the prior probability of its lexicon. The prior of a
lexicon is a function of the complexity of the concepts in its image set. Lexica that use
simpler concepts are \emph{a priori} more likely. This can be motivated by assuming that
learners beam search for suitable concepts to map onto overt signals by (probabilistically)
considering simpler concepts first. There are many ways for defining complexity of a
concept. Following \citet{piantadosi+etal:underreview} and related work, we define complexity
of a concept as a function of its derivation length in a generative ``language of
thought''. For concreteness of example, consider the toy grammar of concepts in
Table~\ref{tab:grammar}. This grammar uses basic set-theoretic operations to form expressions
which can be evaluated as true or false in our three world states. Table~\ref{tab:concepts}
lists the minimal length of a derivation of an expression that represents each concept. A
simple way of forming priors over a lexicon is:
\begin{align*}
  P(L)  & = \prod_{c \in Im(L)} P(c)  \ \ \ \text{, with} & 
  P(c) & \propto \max_{c'}Compl(c') - Compl(c')\,,
\end{align*}
where $Compl(c)$ is the minimum length of a derivation of concept $c$.

>>>>

\begin{table}
  \centering
\begin{center}
  \begin{tabular}{lccclc}
    \toprule
    intuitive name
    & \snone
    & \ssome
    & \sall
    & least complex formula
    & complexity
    \\ \midrule
    ``all''
    & 0
    & 0
    & 1
    & $A \subseteq B$
    & $3$
    \\
    ``some but not all''
    & 0
    & 1
    & 0
    & $A \cap B \neq \emptyset \wedge A \neq \emptyset$
    & $8$
    \\    
    ``some''
    & 0
    & 1
    & 1
    & $A \cap B \neq \emptyset$
    & $4$
    \\
    ``none''
    & 1
    & 0
    & 0
    & $A \cap B = \emptyset$
    & $4$
    \\
    ``none or all''
    & 1
    & 0
    & 1
    & $\neg(A \cap B \neq \emptyset \wedge A \neq \emptyset)$
    & $9$
    \\
    ``not all''
    & 1
    & 1
    & 0
    & $\neg (A \subseteq B)$
    & $5$
    \\
    \bottomrule
  \end{tabular}
\end{center}
\caption{Available concepts and their minimal derivation length}
\label{tab:concepts}
\end{table}

A lexicon $L$ is a mapping $\Messgs \rightarrow C$ from messages to concepts. With three
messages there are $6^3 = 343$ possible lexica, but many of these assign the same concept to
more than one message. For simplicity, we restrict attention to only those lexica which avoid
expressive redundancy. Hardcoding such an \emph{mutual exclusivity bias}
\citep[e.g.][]{Clark2009:Lexical-Meaning}, there are 20 non-redundant lexica. Functional
pressure towards efficient communication will evidently favor non-redundant lexica, so this
restriction is fairly innocuous but practical.

@@

\begin{table}
  \centering
  \begin{align*}
    C & \rightarrow_2 C \wedge C 
    & 
    X & \rightarrow_1 \set{A,B} \\
    C & \rightarrow_2 \neg C 
    & 
    X & \rightarrow_1 X \cap X \\
    C & \rightarrow_1 X \subseteq X
    & 
    X & \rightarrow_1 X \cup X \\
    C & \rightarrow_1 X \neq \emptyset \\
    C & \rightarrow_1 X = \emptyset     
  \end{align*}
  \caption{Toy grammar in a set-theoretic "language of thought"}
  \label{tab:grammar}
\end{table}

The prior probability of a type is just the prior probability of its lexicon. The prior of a
lexicon is a function of the complexity of the concepts in its image set. Lexica that use
simpler concepts are \emph{a priori} more likely. This can be motivated by assuming that
learners beam search for suitable concepts to map onto overt signals by (probabilistically)
considering simpler concepts first. Many ways for defining complexity of a concept are
conceivable. If strong empirical claims were at stake here, empirically motivated measures of
complexity should be used. For the sake of a non-trivial example, we follow
\citet{piantadosi+etal:underreview} and related work to define complexity of a concept as a
function of its derivation cost in a (weighed or probabilistic) generative ``language of
thought''. For concreteness of example, consider the toy grammar of concepts in
Table~\ref{tab:grammar}. This grammar uses basic set-theoretic operations to form expressions
which can be evaluated as true or false in our three world states. Applications of generative
rules have a cost attached to them. (Alternatively, a probability.) Here we simply assume that
Boolean combinations of concepts are more complex than ``atomic'' concepts and that otherwise
each rule application adds the same cost unit. Table~\ref{tab:concepts} lists, for each
concept, the least complex formula derivable in this grammar that has the appropriate truth
conditions. A simple way of defining priors over a lexicon is:
\begin{align*}
  P(L)  & = \prod_{c \in Im(L)} P(c)  \ \ \ \text{, with} & 
  P(c) & \propto \max_{c'}Compl(c') - Compl(c')\,,
\end{align*}
where $Compl(c)$ is the complexity of each concepts.
